Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2019-05-13 unimodular-covers.tex
typos

The main tool is what I call the “parallelepiped circumscribed to a given tetrahedron, defined as follows:

\begin{definition}
Definition: Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$. Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and let $S$ be the convex hull of the $p_i$’s and the $q_i$’s. $S$ is called the \emph{parallelepiped circumscribed} to $T$.
\end{definition}
\paco{define the ``corner tetrahedra''}

@@

1) $S$ is ``half a fundamental domain’’ for the affine lattice $\Lambda_P$ generated by $p_1,…,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique). Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it. (Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).

@@

\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
\paco{rewrite}
4) For every point $p_5 \in S \setminus T$ we have that $\conv(T\cup \{p_5\})$ can be triangulated into three tetrahedra of volume strictly smaller than that of $T$. Proof: $S$ is naturally divided into five tetrahedra: $T$ plus the four ``corner tetrahedra’’ $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$. Suppose wlog that $p_5 \in \conv(p_1,p_2,p_3, q_4)$. Then $\conv(T\cup \{p_5\})$ can be triangulated into 
\[
\conv(p_1,p_2,p_4, p_5),
\conv(p_1,p_3,p_4, p_5),
\conv(p_2,p_3,p_4, p_5).
\]

Putting these properties together we get the following: Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if: ``for every empty tetrahedron $T$ contained in $P$ at least one of the other four vertices of the circumscribed parallelepiped $S$ is also in $P$’'.

@@

\begin{remark}
\label{rem:octahedra_and_tri-prisms}
Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is 
contained in $P$’'. Then, $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.

@@

2’) Even if the answer to (2) is negative, the following can be enough to prove unimodular cover: find a triangulation of $Cayley(P,Q)$ in which all the tetrahedra with an edge in $P$ and an edge in $Q$ have the

>>>>

The main tool is what I call the parallelepiped circumscribed to a given tetrahedron, defined as follows:

\begin{definition}
Definition: Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$. Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and let $S$ be the convex hull of the $p_i$s and the $q_i$s. $S$ is called the \emph{parallelepiped circumscribed} to $T$.
\end{definition}
\paco{define the ``corner tetrahedra''}

@@

1) $S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique). Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it. (Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).

@@

\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
\paco{rewrite}
4) For every point $p_5 \in S \setminus T$ we have that $\conv(T\cup \{p_5\})$ can be triangulated into three tetrahedra of volume strictly smaller than that of $T$. Proof: $S$ is naturally divided into five tetrahedra: $T$ plus the four ``corner tetrahedra'' $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$. Suppose wlog that $p_5 \in \conv(p_1,p_2,p_3, q_4)$. Then $\conv(T\cup \{p_5\})$ can be triangulated into 
\[
\conv(p_1,p_2,p_4, p_5),
\conv(p_1,p_3,p_4, p_5),
\conv(p_2,p_3,p_4, p_5).
\]

Putting these properties together we get the following: Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if: ``for every empty tetrahedron $T$ contained in $P$ at least one of the other four vertices of the circumscribed parallelepiped $S$ is also in $P$''.

@@

\begin{remark}
\label{rem:octahedra_and_tri-prisms}
Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is 
contained in $P$''. Then, $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.

@@

2') Even if the answer to (2) is negative, the following can be enough to prove unimodular cover: find a triangulation of $Cayley(P,Q)$ in which all the tetrahedra with an edge in $P$ and an edge in $Q$ have the
